---
title: "EDA"
author: "Aldrich Paras"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# import data
data <- read.csv("Classification/trainClassification.csv")

```

```{r}
# calculate 5 number summary for each variable
summary(data)
```


```{r}
# Calculate missing counts for each variable
date_missing <-  sum(is.na(data["date"]))
location_missing <- sum(is.na(data["location"]))
min_temp_missing <-  sum(is.na(data["min_temp"]))
max_temp_missing <-  sum(is.na(data["max_temp"]))
rainfall_missing<-  sum(is.na(data["rainfall"]))
evaporation_missing <-  sum(is.na(data["rainfall"]))
sunshine_missing <- sum(is.na(data["sunshine"]))
wind_gust_dir_missing <- sum(is.na(data["wind_gust_dir"]))
wind_gust_speed_missing <- sum(is.na(data["wind_gust_speed"]))
wind_dir9am_missing <- sum(is.na(data["wind_dir9am"]))
wind_dir3pm_missing <- sum(is.na(data["wind_dir3pm"]))
wind_speed9am_missing <- sum(is.na(data["wind_speed9am"]))
wind_speed3pm_missing <- sum(is.na(data["wind_speed3pm"]))
humidity9am_missing <- sum(is.na(data["humidity9am"]))
humidity3pm_missing <- sum(is.na(data["humidity3pm"]))
pressure9am_missing <- sum(is.na(data["pressure9am"]))
pressure3pm_missing <- sum(is.na(data["pressure3pm"]))
cloud9am_missing <- sum(is.na(data["cloud9am"]))
cloud3pm_missing <- sum(is.na(data["cloud3pm"]))
temp9am_missing <- sum(is.na(data["temp9am"]))
temp3pm_missing <- sum(is.na(data["temp3pm"]))
rain_today_missing <- sum(is.na(data["rain_today"]))
rain_tomorrow_missing <- sum(is.na(data["rain_tomorrow"]))

# percentage of missing values
missing_percent <- mean(is.na(data))
missing_percent
```
date - 0 Missing
location - 0 Missing
min_temp - 157 Missing
max_temp - 70 Missing
rainfall - 357 Missing
evaporation - 357 Missing
sunshine - 33566 Missing
wind_gust_dir - 2285 Missing
wind_gust_speed - 2271 Missing
wind_dir9am - 2432 Missing
wind_dir3pm -  941 Missing
wind_speed9am - 328 Missing
wind_speed3pm - 650 Missing
humidity9am - 453 Missing
humidity3pm - 897 Missing
pressure9am - 3381 Missing
pressure3pm - 3376 Missing
cloud9am - 12886 Missing
cloud3pm - 13688 Missing
temp9am - 233 Missing
temp3pm - 666 Missing
rain_today - 357 Missing
rain_tomorrow - 0 Missing

~13.74% of data is missing

```{r}
# count information and graphs for variables
library(ggplot2)

# Location counts
ggplot(data, aes(x = location))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip()+ggtitle("Location")

# min_temp counts
ggplot(data, aes(x = min_temp))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip()+ggtitle("Min Temp")

# max_temp counts
ggplot(data, aes(x = max_temp))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip()+ggtitle("Max Temp")

# rainfall counts
ggplot(data, aes(x = rainfall))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Rainfall") + xlim(0,15) + ylim(0,100)

# evaporation counts
ggplot(data, aes(x = evaporation))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Evaporation") 

# sunshine counts
ggplot(data, aes(x = sunshine))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Sunshine") 

# wind gust direction counts
ggplot(data, aes(x = wind_gust_dir))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Wind Gust Direction") 

# Wind Gust Speed counts
ggplot(data, aes(x = wind_gust_speed))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Wind Gust Speed") 

# Wind Direction 9 am
ggplot(data, aes(x = wind_dir9am))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Wind Direction 9 am") 

# Wind Direction3 pm
ggplot(data, aes(x = wind_dir3pm))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Wind Direction 3 pm") 

# Wind Speed 9 am
ggplot(data, aes(x = wind_speed9am))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Wind Speed 9 am") 

# Wind Speed 3 pm
ggplot(data, aes(x = wind_speed3pm))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Wind Speed 3 pm") 

# Humidity 9 am
ggplot(data, aes(x = humidity9am))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Humidity 9 am") 

# Humidity 3 pm
ggplot(data, aes(x = humidity3pm))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Humidity 3 pm") 

# Pressure 9 am
ggplot(data, aes(x = pressure9am))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Pressure 9 am") 

# Pressure 3 pm
ggplot(data, aes(x = pressure3pm))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Pressure 3 pm") 

# Cloud 9 am
ggplot(data, aes(x = cloud9am))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Cloud 9 am") 

# Cloud 3 pm
ggplot(data, aes(x = cloud3pm))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Cloud 3 pm")

# Temp 9 am
ggplot(data, aes(x = temp9am))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Temp 9 am")

# Temp 3 pm
ggplot(data, aes(x = temp3pm))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Temp 3 pm")

# rain today
ggplot(data, aes(x = rain_today))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Rain Today")

# rain tomorrow
ggplot(data, aes(x = rain_tomorrow))+ geom_histogram(stat = "count", fill = "tomato2")+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + coord_flip() +ggtitle("Rain Tomorrow")
```


